Worflow

### 2017-06-12 R/workflow.Rmd
update zinbFit for slingshot

```{r options, echo=FALSE, results="hide",message=FALSE, error=FALSE, include=FALSE, autodep=TRUE}
knitr::opts_chunk$set(fig.align="center", cache=TRUE, error=FALSE, message=FALSE, warning=TRUE)
#bioconductor:
library(scRNAseq)
library(scone)
library(clusterExperiment)
library(BiocParallel)
#github
library(zinbwave)
library(slingshot)
#CRAN
library(doParallel)
library(ggplot2)
library(gplots)
library(magrittr)
library(matrixStats)
library(Rtsne)
library(RColorBrewer)
library(digest)
library(rARPACK)
set.seed(20)
if(packageVersion("clusterExperiment")<'1.3.0.9008') stop("must have current develop version to avoid bugs")
if(packageVersion("zinbwave")<'0.99.4.2') stop("must have current develop version to avoid bugs")
```

@@

FP: for zinbwave, parallel computing is now handle by BiocParallel. If you have a Windows machine, please update the code to allow paralell computing.

@@

FP: Elizabeth, I could not install the last version 1.3.0.9009 of clusterExperiment. The error is

checking DESCRIPTION meta-information ... ERROR
Malformed Depends or Suggests or Imports or Enhances field.
Offending entries:
  MAST
BiocStyle
Entries must be names of packages optionally followed by '<=' or '>=',
white space, and a valid version number in parentheses.

@@

```{r zinb}
fn <- '../data/zinb_batch.rda'
if (runZinb & !file.exists(fn)){
  print(system.time(zinb <- zinbDimRed(se, K = 50, X = '~ batch',
                                       residuals = T,
                                       normalizedValues = T)))
  save(zinb, file = fn)
}else{
  load(fn)
}
```

@@

```{r norm}
res <- assays(zinb)[['normalizedValues']]
res[1:3,1:3]
```

@@

```{r lookatnorm}
res_order <- res[, order(as.numeric(batch))]
col_order <- as.numeric(batch)[order(as.numeric(batch))]
boxplot(res_order, main='Boxplot of normalized values\ncolor=batch',
        col = col_order, staplewex = 0, outline = 0, border = col_order, xaxt = 'n')
```

@@

```{r pcanorm}
pca <- prcomp(t(res))
par(mfrow = c(1,2))
plot(pca$x, col = col_batch[batch], pch = 20,
     main="PCA of normalized values\ncolor=batch")
plot(pca$x, col = col_clus[as.character(clus.labels)], pch = 20,
     main = "PCA of normalized values\ncolor=cluster")
par(mfrow = c(1,1))
```

@@

```{r rsec_50}
fn <- '../data/RSEC_W.rda'
if (runClus & !file.exists(fn)){
  #symbol for samples missing from original clustering
	metadata$clusterLabels[is.na(metadata$clusterLabels)]<- -2 
  seObj<-SummarizedExperiment(t(W),colData=metadata)
  print(system.time(ceObj <- RSEC(seObj, k0s = 4:15, alphas = c(0.1),
                                  betas = 0.8,
                clusterFunction = "hierarchical01", minSizes=1,
                ncores = NCORES, isCount=FALSE,
                subsampleArgs = list(resamp.num=100,
                                     clusterFunction="kmeans",
                                     clusterArgs=list(nstart=10)),
                seqArgs = list(k.min=3, top.can=5), verbose=TRUE,
                combineProportion = 0.7,
                mergeMethod = "none")))
  save(ceObj, file = fn)
}else{
  load(fn)
}
```

@@

plotHeatmap(ceObj, visualizeData = assays(se)$normalizedValues,
            whichClusters = "primary",clusterFeaturesData="all",
            clusterSamplesData = "dendrogramValue",
			sampleData=c("clusterLabels","Batch","Experiment"), clusterLegend=metaColors, annLegend=FALSE,
            main = 'Normalized values, 1000 most variable genes',
            breaks = 0.99)

```

@@

Cluster name | Description | Correspondence | Color
-------------|-------------|-----------------|---------
c1 | HBC | original 1, 5 | darkblue
c2 | new and small | new and small | green
c3 | new and small | new and small | red
c4 | GBC / immature neurons / MV 1 | original 2, 3, 11, 14 | orange
c5 | Sus | original 4, 7 | darkpurple
c6 | Neuron | original 9, 12 | brown
c7 | Immature Neuron | original 10, 14 | cian 
c8 | Immature Neuron | original 14 | lightpurple
c9 | Microvillus | original 15 | grayblue

FP: We need to work on the cluster coloring (!) and the matching with the clusters you found for the paper with Russell, maybe by changing the parameters in clusterMany?

EP: The function plotClusters aligns the clusters and can return the aligned coloring between clusterings. There is an option to return an updated CE object that changes the colors in clusterLegend to match the new colors found by aligning the clusters (`resetNames` and `resetColors` arguments). These haven't been extensively tested in real life because I don't think we've used them much. But it would be good to demonstrate!

FP: With all the improvements made, I think we simply have the correct colors now.

@@

  lineages <- get_lineages(X, clus.labels = cl, start.clus = "c1",
                           end.clus = c("c5", "c6", "c9"))
  curves <- get_curves(X, clus.labels = cl, lineages = lineages)
  plot_curves(X, cl, curves, col.clus = pal)
  plot_tree(X, cl, lineages, col.clus = pal)

@@

```{r refit_zinb}
mod <- model.matrix(~ batch[our_cl != "-1"])
fn <- '../data/refit_zinbwave_slingshot.rda'
if (runZinb & !file.exists(fn)){
  zinbList <- lapply(Kvec, function(k){
    zinbFit(core[, our_cl != "-1"], X = mod, K = k, ncores = NCORES)
  })
  save(zinbList, file = fn)
}else{
  load(fn)
}
```

>>>>

```{r options, echo=FALSE, results="hide",message=FALSE, error=FALSE, include=FALSE, autodep=TRUE}
knitr::opts_chunk$set(fig.align="center", cache=TRUE, error=FALSE, message=FALSE, warning=TRUE)
#bioconductor:
library(scRNAseq)
library(scone)
library(clusterExperiment)
library(BiocParallel)
#github
library(zinbwave)
library(slingshot)
#CRAN
library(doParallel)
library(ggplot2)
library(gplots)
library(magrittr)
library(matrixStats)
library(Rtsne)
library(RColorBrewer)
library(digest)
library(rARPACK)
set.seed(20)
if(packageVersion("clusterExperiment")<'1.3.0.9009') stop("must have current develop version to avoid bugs")
if(packageVersion("zinbwave")<'0.99.4.2') stop("must have current develop version to avoid bugs")
```

@@

FP: Version of zinbwave to use is last version of branch normvalues (branched from master). Parallel computing is now handle by BiocParallel. If you have a Windows machine, please update the code to allow paralell computing.

@@

@@

```{r zinb}
fn <- '../data/zinb_batch.rda'
if (runZinb & !file.exists(fn)){
  print(system.time(se <- zinbDimRed(se, K = 50, X = '~ Batch',
                                       residuals = T,
                                       normalizedValues = T)))
  save(se, file = fn)
}else{
  load(fn)
}
```

@@

FP: note to myself: why do we have infinite values in the residuals now? It shows the same results as before, but we should not see infinite values here!

```{r norm}
norm <- assays(se)$normalizedValues
if (sum(is.infinite(norm))>0){
  maxNorm = max(norm[!is.infinite(norm)])
  assays(se)$normalizedValues[is.infinite(norm)] <- maxNorm
  norm <- assays(se)$normalizedValues
}
norm[1:3,1:3]
```

@@

```{r lookatnorm}
norm_order <- norm[, order(as.numeric(batch))]
col_order <- as.numeric(batch)[order(as.numeric(batch))]
boxplot(norm_order, main='Boxplot of normalized values\ncolor=batch',
        col = col_order, staplewex = 0, outline = 0, border = col_order,
        xaxt = 'n')
```

@@

```{r pcanorm}
pca <- prcomp(t(norm))
par(mfrow = c(1,2))
plot(pca$x, col = col_batch[batch], pch = 20,
     main="PCA of normalized values\ncolor=batch")
plot(pca$x, col = col_clus[as.character(clus.labels)], pch = 20,
     main = "PCA of normalized values\ncolor=cluster")
par(mfrow = c(1,1))
```

@@

```{r rsec_50}
fn <- '../data/RSEC_W.rda'
if (runClus & !file.exists(fn)){
  #symbol for samples missing from original clustering
	metadata$clusterLabels[is.na(metadata$clusterLabels)] <- -2 
  seObj <- SummarizedExperiment(t(W), colData = metadata)
  print(system.time(ceObj <- RSEC(seObj, k0s = 4:15, alphas = c(0.1),
                                  betas = 0.8,
                clusterFunction = "hierarchical01", minSizes=1,
                ncores = NCORES, isCount=FALSE,
                subsampleArgs = list(resamp.num=100,
                                     clusterFunction="kmeans",
                                     clusterArgs=list(nstart=10)),
                seqArgs = list(k.min=3, top.can=5), verbose=TRUE,
                combineProportion = 0.7,
                mergeMethod = "none")))
  save(ceObj, file = fn)
}else{
  load(fn)
}
```

@@

plotHeatmap(ceObj, visualizeData = assays(se)$normalizedValues,
            whichClusters = "primary",clusterFeaturesData="all",
            clusterSamplesData = "dendrogramValue",
            sampleData=c("clusterLabels","Batch","Experiment"),
            clusterLegend=metaColors, annLegend=FALSE,
            main = 'Normalized values, 1000 most variable genes',
            breaks = 0.99)
```

@@

Cluster name | Description | Correspondence
-------------|-------------|-----------------
c1 | HBC | original 1, 5
c2 | new and small | new and small
c3 | new and small | new and small
c4 | GBC / immature neurons / MV 1 | original 2, 3, 11, 14
c5 | Sus | original 4, 7
c6 | Neuron | original 9, 12
c7 | Immature Neuron | original 10, 14
c8 | Immature Neuron | original 14
c9 | Microvillus | original 15

@@

  lineages <- get_lineages(X, clus.labels = cl, start.clus = "c1",
                           end.clus = c("c3", "c6", "c10"))
  curves <- get_curves(X, clus.labels = cl, lineages = lineages)
  plot_curves(X, cl, curves, col.clus = pal)
  plot_tree(X, cl, lineages, col.clus = pal)

@@

```{r refit_zinb}
fn <- '../data/refit_zinbwave_slingshot.rda'
if (runZinb & !file.exists(fn)){
  zinbList <- lapply(Kvec, function(k){
    zinbFit(se[, our_cl != "-1"], X = '~ Batch', K = k)
  })
  save(zinbList, file = fn)
}else{
  load(fn)
}
```
